---
title: "hw3"
author: "Enbo Tian"
date: "2022/2/28"
output:
  word_document: default
  html_document: default
  pdf_document: default
---
# Graphical Model 1
## a)
```{r}
rm(list = ls())
X <- c("cold", "hot", "mild")
# day 0
day0 = replicate(5,sample(X,size = 1,prob=c(1/3,1/3,1/3)))
# function of day
dayk <- function(day){
  dayk <- rep(0,5)
  for (i in 1:5){
    if (day[i] == "cold"){
    dayk[i] = sample(c("cold", "hot", "mild"),size =1,prob = c(1/2,1/4,1/4))
    }
    else if (day[i] == "hot"){
    dayk[i] = sample(c("cold", "hot", "mild"),size =1,prob = c(1/3,1/3,1/3))
    }
    else if(day[i] == "mild"){
      dayk[i]= sample(c("cold", "hot", "mild"),size =1,prob = c(1/4,1/4,1/2))
    }
  }
  dayk
}
# day 1:5
day1 <- dayk(day0)
day2 <- dayk(day1)
day3 <- dayk(day2)
day4 <- dayk(day3)

day <- data.frame(day0, day1, day2, day3, day4)
day
```

## b)
```{r}
# P(day0)
p0 <- c(1/3,1/3,1/3)
# p (k given k-1)
pgiven <- matrix(c(1/2,1/4,1/4,1/3,1/3,1/3,1/4,1/4,1/2),ncol=3)
# marginal prob
p1 <- pgiven%*%p0
p2 <- pgiven%*%p1
p3 <- pgiven%*%p2
margp <- data.frame(p0,p1,p2,p3)
margp
```

## c)
```{r}
# 3|2hot 
p3g2 <- c(1/3,1/3,1/3)
p3g2
# p(1|2="hot") = p(2|1)*p(1)/p(2 = "hot")
p1 <- c(p1)
p2 <- c(p2)
p3 <- c(p3)
p1g2 <- pgiven * p1 / p2 
p1g2 <- c(p1g2[,2])
p1g2
# p(0|1) = p(1|0)*p(0)/p(1|2="hot")
p0g1 <- pgiven * p0 /p1g2
p0g1
```

## d)
```{r}
# give day2 is hot
day2 = "hot"
# get most probable day1
if (max(p1g2) == p1g2[1]){
  day1 = "cold"
  i = 1
}else if(max(p1g2) == p1g2[2]){
  day1 = "hot"
  i = 2
}else if(max(p1g2) == p1g2[3]){
  day1 = "mild"
  i =3
}
# get most probable day0
new_p0g1 <- p0g1[,i]
if (max(new_p0g1) == new_p0g1[1]){
  day0 = "cold"
}else if(max(new_p0g1) == new_p0g1[2]){
  day0 = "hot"
}else if(max(new_p0g1) == new_p0g1[3]){
  day0 = "mild"
}
# Same prob for day3 given hot of day2
c(day0,day1,day2)
```
 the most probable report for day 0 to 2 are "cold" "hot"  "hot", and 
 
 we have the same probability for day3.

# Graphical Model 2
## a)
```{r}
mi <- c(-2, 2 ,0)
# height function
height <- function(statep){
  m <- replicate(5,sample(mi,size = 1,prob = statep))
  y <- rep(0,5)
  for (i in 1:5){
    y[i] <- rnorm(1,m,1)
  }
  y
}
# get height
y0 <- height(p0)
y1 <- height(p1)
y2 <- height(p2)
y3 <- height(p3)
datay <- data.frame(y0,y1,y2,y3)
datay
```

## b)
```{r}
m <- c(2,0,-2,-2)
y0 <-replicate(5,rnorm(1,m[1],1)) 
y1 <-replicate(5,rnorm(1,m[2],1))
y2 <-replicate(5,rnorm(1,m[3],1))
y3 <-replicate(5,rnorm(1,m[4],1))
data2y<- data.frame(y0,y1,y2,y3)
data2y
```
## c)
```{r}
Y0 =0.7
Y1 =1.5
Y2 =-1.8
Y3 =-1

#p(y0|x0)
d0 <- rep(0,3)
d0[1] <- dnorm(Y0,-2,1)
d0[2] <- dnorm(Y0,0,1)
d0[3] <- dnorm(Y0,2,1)
#p(y1|x1)
d1 <- rep(0,3)
d1[1] <- dnorm(Y1,-2,1)
d1[2] <- dnorm(Y1,0,1)
d1[3] <- dnorm(Y1,2,1)
#p(y2|x2)
d2 <- rep(0,3)
d2[1] <- dnorm(Y2,-2,1)
d2[2] <- dnorm(Y2,0,1)
d2[3] <- dnorm(Y2,2,1)
#p(y1|x1)
d3 <- rep(0,3)
d3[1] <- dnorm(Y3,-2,1)
d3[2] <- dnorm(Y3,0,1)
d3[3] <- dnorm(Y3,2,1)
# p(x0)*p(y0|x0)*p(x1|x0)*p(y1|x1)*p(x2|x1)*p(y2|x2)*p(x3|x2)*p(y3|x3)
mp <- p0*d0*pgiven*d1*pgiven*d2*pgiven*d3
sum(mp[,1])
sum(mp[,2])
sum(mp[,3])
```
# Gaussian Mixture Model
## a)
```{r}
rm(list = ls())
library("readxl")
data <- read_excel("gmm_data.xlsx")
plot(data$Var1,data$Var2)
```
the number of clusters is 3.

## b)
```{r}
library(fMultivar)
msnFit(data)

library(ellipse)
rho = cor(data)
y_on_x <- lm(data$Var2 ~ data$Var1)
x_on_y <- lm(data$Var1 ~ data$Var2)
plot_legend <- c("99% CI green", "95% CI red","90% CI blue",
                 "Y on X black", "X on Y brown")
plot(data, xlab = "X", ylab = "Y",col = "grey")
lines(ellipse(rho), col="red")
lines(ellipse(rho, level = .99), col="green")
lines(ellipse(rho, level = .90), col="blue")
abline(y_on_x)
abline(x_on_y, col="brown")
legend(3,1,legend=plot_legend,cex = .5, bty = "n")
```
## c)
```{r}
library(MGMM)
d <- as.matrix(data)
K2GMM <- FitGMM(d,k=2)
plot(data, xlab = "X", ylab = "Y",col = "grey")
points(K2GMM@Means[[1]][1],K2GMM@Means[[1]][2],col = "red")
points(K2GMM@Means[[2]][1],K2GMM@Means[[2]][2],col = "dark green")
```
## d)
```{r}
K3GMM <- FitGMM(d,k=3)
plot(data, xlab = "X", ylab = "Y",col = "grey")
points(K3GMM@Means[[1]][1],K3GMM@Means[[1]][2],col = "red")
points(K3GMM@Means[[2]][1],K3GMM@Means[[2]][2],col = "dark green")
points(K3GMM@Means[[3]][1],K3GMM@Means[[3]][2],col = "blue")
```
## e)

```{r}
# Sigma_i = dx * d covariance matrix
si <- diff(data$Var1) %*% diff(rho)
si[1:20,]
```
# Poisson Mixture Model
## 1)
```{r}
rm(list = ls())
```
```{r}
library("readxl")
data <- read_excel("poisson_data.xlsx")
x <- data$X
hist(x)
```
From the plot it may fit two distribution from 1-10 and 10-30

## b)
```{r}
library(MASS)
lambda <- fitdistr(x,densfun="Poisson")
lambda
a <- 0:30
b <- dpois(a,lambda$estimate)
hist(x,xlim=c(0,30),ylim= c(0,250))
par(new=TRUE)
plot(a,b,yaxt="n",xaxt="n",xlab="",ylab="")
```
a simple poisson distribution is not a good fit.

## c)
```{r}
library(mixtools)
mixture <- normalmixEM(x,lambda = 0.092997,k=2)
summary(mixture)
```

## d)
```{r}
plot(mixture, density=TRUE)
```

## e)
the single poisson distribution can not fit the data very well,
the mixture model give two poisson distribution and separate the data into 
two modle, which fit the data better. 