---
title: "cs539 hw5"
author: "Enbo Tian"
date: '2022-04-19'
output:
  word_document: default
  html_document: default
  pdf_document: default
---
#KL Distance 

## a)
```{r}
library(LaplacesDemon)
p <- 1/3*dnorm(runif(10),-1,2)+2/3*dnorm(runif(10),1,1)

m <- 1:20/20*2-1 # set 100 m from -1 to 1
s <- 1:10/10+1  # set 100 s from 1 to 2

Dist <- rep(0,200)
count = 1
for(i in 1:20){
  for(j in 1:10){
    q <-  dnorm(runif(10),m[i],s[j])
    Dist[count] <- KLD(p,q)$intrinsic.discrepancy
    count = count+1
  }
}

Dist <- matrix(Dist,20,10)
Dist
min(Dist)
which(Dist== min(Dist), arr.ind = TRUE)

# row = 8, col= 3
8/20-1
3/10+1
```
Then $m_a = -0.6$ , $\sigma^2_m = 1.3$ have the minimum KL distance with p 

## b)
```{r}
mean(p)
var(p)
q_p <- dnorm(runif(10),0.2809131,0.0009584375)

KLD(p,q)
```
the KL distanced is 0.004042311, however, it is not the minimum KL distance.

# TSNE

## a)
```{r}
library(Rtsne)
xtrain <- read.table("X_train.txt")

ytrain <- read.table("y_train.txt")

traintsne <- Rtsne(xtrain)
# Y
# 1   red WALKING
# 2   green WALKING_UPSTAIRS
# 3   blue WALKING_DOWNSTAIRS
# 4   light_blue SITTING
# 5   purple STANDING
# 6   yellow LAYING
plot(traintsne$Y,col= ytrain$V1+1)
```
The KL distance for Y=1 is the largest, Y = 7 is the smallest, Y = 4,5 are emed.


## b)
```{r}
xtest <- read.table("X_test.txt")
ytest <- read.table("y_test.txt")
testtsne <- Rtsne(xtest)
plot(testtsne$Y,col= ytest$V1+1)
```
The KL distance for red, green, blue(1, 2, 3) are large, 4,5,6 are small. 

## c)
The  similarities part are the each part distribution of the color groups.
both 4 and 5 are neighboring, 1 and 2 are dispersive.

# Neural Networks
## a)
```{r}
library(neuralnet)
Y1 <- ytrain$V1
training <- data.frame(Y1,xtrain)
TrainNN <- neuralnet(Y1 ~ .,data = training)
summary(TrainNN)
plot(TrainNN,col.hidden = 'darkgreen',     
col.hidden.synapse = 'darkgreen',
     show.weights = F,
     information = F,
     fill = 'lightblue')
```
```{r}
Y2 <- ytest$V1
testing <- data.frame(Y2,xtest)
TestNN <- neuralnet(Y2 ~ .,data = testing)
summary(TestNN)
plot(TestNN,col.hidden = 'darkgreen',     
col.hidden.synapse = 'darkgreen',
     show.weights = F,
     information = F,
     fill = 'lightblue')

```

## b)
```{r}
trainpartci<-read.table("subject_train.txt")
testpartci <- read.table("subject_test.txt")

Y3 <- trainpartci$V1
training <- data.frame(Y3,xtrain)
TrainNN <- neuralnet(Y3 ~ .,data = training)
plot(TrainNN,col.hidden = 'darkgreen',     
col.hidden.synapse = 'darkgreen',
     show.weights = F,
     information = F,
     fill = 'lightblue')
```
```{r}
Y4 <- testpartci$V1
testing <- data.frame(Y4,xtest)
TestNN <- neuralnet(Y4 ~ .,data = testing)
plot(TestNN,col.hidden = 'darkgreen',     
col.hidden.synapse = 'darkgreen',
     show.weights = F,
     information = F,
     fill = 'lightblue')

```
